Random Forests

### 2020-07-01 pkg/partykit/inst/ULGcourse-2020/slides_forest.Rnw
small revisions and formatting
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@3027 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 2c59e2eda706c3cec5c169f8a7d9832bf0c30e67

\documentclass[11pt,t,usepdftitle=false,aspectratio=169]{beamer}
\usetheme[nototalframenumber,url,license]{uibk}

\title{Random Forests}
\author{Achim Zeileis, Lisa Schlosser}
\setbeamerfont{url}{size*={11.5pt}{13pt},series=\mdseries}
%\URL{http://www.partykit.org/partykit/}
\renewcommand{\headerimage}[1]{%
   \IfStrEqCase{#1}{%
      {1}{%
         \gdef\myheaderimageid{#1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{forest.jpg}%
      }}[%
         \gdef\myheaderimageid{1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{forest.jpg}%
      ]%
}
\headerimage{1}

@@

\setbeamertemplate{caption}{\insertcaption} 
%% includes a replacement for \usepackage{Sweave}
%\usepackage{Sweave}
\usepackage{changepage}
\usepackage{amsmath,tikz}
\usepackage{calc}
\usepackage{graphicx}
\usetikzlibrary{positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap,trees,tikzmark,decorations.pathreplacing}
\usepackage{xcolor}
%\usepackage[cal=boondoxo]{mathalfa}
%\graphicspath{{plots/}}
\newcommand{\argmin}{\operatorname{argmin}\displaylimits}

@@

\section{Forests models}

\subsection{Motivation: Forest models}
\begin{frame}[fragile]
\frametitle{Motivation: Forest models}

@@

\begin{frame}
\frametitle{Motivation: Forest models}

@@

\subsection{Aggregation of trees}

@@

The predicted response $\hat{y}$ or, more generally, the predicted model parameter (vector) $\hat{\theta}$ is the argument of the parameter space $\Theta$ that minimizes the weighted sum of the objective function $f$ evaluated for learning data $\{y_i\}_{i=1,\ldots,n}$:

@@

\[
\hat{\theta}(x) =  \argmin_{\theta \in \Theta} \sum_{i=1}^n w_i(x) \cdot f(\theta; y_i)
\]

@@

\only<5->{
%\vspace{-3cm}
{\bf Note:} This general notation holds for classic random forests averaging over mean values of tree nodes by setting $f(\theta, y_i) = (\theta - y_i)^2$\\
but also allows for combining quantiles (quantile regression forests) or full distributions as in distributional forests with
$f(\theta, y_i) = log(likelihood(\theta, y_i))$.\\
}

@@

\subsection{Subsampling}

@@

<<call_randomForest, eval=FALSE, echo=TRUE>>=
randomForest(formula, data=NULL, ..., subset, na.action=na.fail)
randomForest(x, y=NULL,  ...)
@

Optional control arguments:
\begin{itemize}
\item \code{ntree}: nr. of trees, 
\item \code{mtry}: nr. of randomly sampled split variable candidates, 
\item \code{replace}: type of subsetting, 
\item \code{sampsize}: size of drawn subsamples, 
\item \code{nodesize}: minimum size of terminal nodes, 
\item \code{maxnodes}: maximum number of terminal nodes, 
\item \ldots 
\end{itemize}

@@

@@

\textbf{Further R packages:}
%
\begin{itemize}
  \item ranger (classic random forests),
  \item grf (generalized random forests),
  \item disttree (distributional forests, based on partykit),
  \item \ldots
\end{itemize}

@@

\subsection{References}

>>>>

\documentclass[11pt,t,usepdftitle=false,aspectratio=169]{beamer}
\usetheme[nototalframenumber,license]{uibk}

\title{Random Forests}
\subtitle{Supervised Learning: Algorithmic Modeling}
\author{Lisa Schlosser, Achim Zeileis}

%% forest header image
\renewcommand{\headerimage}[1]{%
   \IfStrEqCase{#1}{%
      {1}{%
         \gdef\myheaderimageid{#1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{forest.jpg}%
      }}[%
         \gdef\myheaderimageid{1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{forest.jpg}%
      ]%
}
\headerimage{1}

%% custom subsection slides
\setbeamercolor*{subsectionfade}{use={normal text},parent={normal text},fg=structure.fg!30!normal text.bg}
\AtBeginSubsection[]{%
  \begin{frame}[c]
    \begin{center}
      \usebeamercolor[fg]{subsectionfade}
      \Large \insertsection \\[2ex]
      \usebeamercolor[fg]{structure}
      \huge\bfseries\insertsubsection
    \end{center}
  \end{frame}
}

@@

\setbeamertemplate{caption}{\insertcaption} 
%% includes a replacement for \usepackage{Sweave}
%\usepackage{Sweave}
\usepackage{changepage}
\usepackage{amsmath,tikz}
\usepackage{calc}
\usepackage{graphicx}
\usetikzlibrary{positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap,trees,tikzmark,decorations.pathreplacing}
\usepackage{xcolor}
\newcommand{\argmin}{\operatorname{argmin}\displaylimits}

@@

\section{Random forests}

\subsection{Motivation}
\begin{frame}[fragile]
\frametitle{Motivation}

@@

\begin{frame}
\frametitle{Motivation}

@@

\subsection{Aggregation of trees\\
\&\\
supsampling}

@@

The predicted response $\hat{y}$ 
%or, more generally, the predicted model parameter (vector) $\hat{\theta}$ 
is the argument %of the parameter space $\Theta$ 
that minimizes the weighted sum of the objective function $f$ evaluated for learning data $\{y_i\}_{i=1,\ldots,n}$:

@@

\[
\hat{y}(x) =  \argmin_{y} \sum_{i=1}^n w_i(x) \cdot f(y; y_i)
\]

@@

\only<5->{
%\vspace{-3cm}
{\bf Note:} In model-based trees and forests a prediction is made not only for the response but for a model parameter (vector) $\theta$. 
This more general approach holds for classic random forests averaging over mean values of tree nodes by setting $f(\theta, y_i) = (\theta - y_i)^2$\\
but also allows for combining quantiles (quantile regression forests) or full distributions as in distributional forests with
$f(\theta, y_i) = log(likelihood(\theta, y_i))$.\\
}

@@

@@

% \medskip
% 
% <<call_randomForest, eval=FALSE, echo=TRUE>>=
% randomForest(formula, data=NULL, ..., subset, na.action=na.fail)
% randomForest(x, y=NULL,  ...)
% @
% 
% \medskip
% 
% Optional control arguments:
% \begin{itemize}
% \item \code{ntree}: nr. of trees, 
% \item \code{mtry}: nr. of randomly sampled split variable candidates, 
% \item \code{replace}: type of subsetting, 
% \item \code{sampsize}: size of drawn subsamples, 
% \item \code{nodesize}: minimum size of terminal nodes, 
% \item \code{maxnodes}: maximum number of terminal nodes, 
% \item \ldots 
% \end{itemize}

@@

{\bf ranger}:
\begin{itemize}
\item RANdom  forest  GEneRator.
\item R package providing a faster implementation of classic random forests.
\end{itemize}

<<call_ranger, eval=FALSE, echo=TRUE>>=
ranger(formula, data)
@

 Optional control arguments:
\begin{itemize}
\item \code{num.trees}: number of trees, 
\item \code{mtry}: number of randomly sampled split variable candidates, 
\item \code{replace}: sample with replacement, 
\item \code{sample.fraction}: fraction of observations to sample, 
\item \code{min.node.size}: minimal size of terminal nodes, 
\item \code{max.depth}: maximal tree depth, 
\item \code{splitrule}: splitting rule, 
\item \ldots 
\end{itemize}
\end{frame}

@@

\textbf{Further R packages:}
%
\begin{itemize}
  \item grf (generalized random forests),
  \item disttree (distributional forests, based on partykit),
  \item \ldots
\end{itemize}

@@

%\subsection{References}
